---
title: "Relationship between Party Performance and Constituency in Maharashtra"
author: "Tanvi Dani"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [Prez2.css, metropolis, metropolis-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, echo = FALSE, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  dev = 'png',
  fig.width = 8, fig.asp=0.80, fig.align= 'center', out.width= "80%")
```
``` {r, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)


library(knitr)
library(readxl)
library(ggplot2)
library(kableExtra)

df = read_csv('/Users/tanvidani/Library/CloudStorage/OneDrive-american.edu/Data Analysis/TCPD_Mah.csv')
```
``` {r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, results - "hide"}

#convert into factor

# fct_count(df$MyNeta_education, prop = T)
df$MyNeta_education = as_factor(df$MyNeta_education)

df1 = df%>%
  mutate(
    Education_Level = fct_collapse(MyNeta_education,
                                   Minimal = c('Illiterate', 'Literate', '5th Pass'),
                                   HighSchool = c('8th Pass','10th Pass', '12th Pass'),
                                   College = c('Graduate','Graduate Professional'),
                                   Post_Grad = c('Post Graduate', 'Doctorate'),
                                   Other = 'Others'
    )
  )
```


# Trivedi Center for Political Data

### Comparing Party performance based on Constituencies


- Trivedi Center for Political data- a research center at Ashoka University in India. 

- Data on incumbents in the Individual Incumbency Data set

- Presentation on data filtered for the state of Maharashtra and 2014, 2019 General Elections

- Data has 8913 observations. 

- Variables considered- Education Level, Constituency type

``` {r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
df1 %>%
  count(Constituency_Type) %>%
  kbl(.,align = rep("c",5), digits = 1L) %>%
    kable_classic(full_width = F, html_font = "Cambria")
```


---

## Variable 1- Education level of Candidates

``` {r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.asp= 0.60}
df1%>%
  ggplot(aes(x = Education_Level)) +
  geom_bar(fill = 'steelblue2') +
  scale_x_discrete(na.translate = F) +
  scale_y_continuous(expand = c(0,0)) +
  theme_minimal() + 
   labs(x = "Education level of candidates")
```


---

## Comparing Educational level across 3 constituencies

### Elections considered: 

- 2019 General Election
- 2014 General Election


### Constituency type

  - GEN- General 
  
  - SC- Scheduled Caste
  
  - ST- Scheduled Tribe
  
  
- Each candidate contests general elections from their specific reserved constituencies. All constituencies in a state are reserved for candidates within these 3 categories.  


---

## Mapping educational levels across constituencies


``` {r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.width= 12, fig.asp=  0.60}
df1 %>%
  drop_na(Constituency_Type, Education_Level) %>%
  ggplot(aes(x = Education_Level, fill = Constituency_Type)) +
  geom_bar(position= 'dodge') +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1000), breaks = c(0, 200, 400, 600, 800, 1000)) +
  theme_classic() +
  labs(x = "Education level of candidates")

```

---

## Observations 

- Most candidates with high school and college level education are from the General Constituency
- A significant number of SC constituency candidates have High School and College level education
- Very few candidates from any of the constituencies have post-graduate education

.pull-right[
``` {r, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE, fig.asp= 0.60}
df1 %>%
  group_by(Education_Level) %>%
  count(Constituency_Type) %>%
  pivot_wider(
    names_from = Constituency_Type,
    values_from = n
  ) %>%
  kbl(.,align = rep("c",5), digits = 1L) %>%
    kable_classic(full_width = F, html_font = "Cambria")
```
]


---

class: center, middle

## Conclusion 

### Unlike common wisdom of candidates in General elections being mostly uneducated, we can see that a considerable number of candidates, from each category have had high school, college, and even post-graduate education. Candidates with minimal education are only 346 out of 8912 candidates from all 3 categories considered, which is 3.9%. The numbers for college graduates contensting elections can still be improved. 